Do investors in clean energy ETFs herd?

### 2024-12-09 alt_energy_herding.tex
cleaning up references

\author { 
Vassilios Babalos\footnote{Department of Accounting and Finance, University of Peloponnese, Kalamata,  Greece; Email: v.babalos@uop.gr.} \,\,
Xolani Sibande\footnote{Corresponding Author. Department of Economics, University of Pretoria, Pretoria, South Africa; Email: xolaniss@gmail.com. Economic Research Department, South African Reserve Bank, Pretoria, South Africa.} \,\, 
% Riza Demirer\footnote{Department of Economics and Finance, Southern Illinois University Edwardsville, Edwardsville, IL 62026-1102, USA; Email: rdemire@siue.edu} \,\,
Rangan Gupta\footnote{Department of Economics, University of Pretoria, Pretoria, South Africa; Email: rangan.gupta@up.ac.za.}
}
\date{\today}
\maketitle

@@

Argument: Investment interest on green ETFs has been explosive mainly by
climate risks faced by the global economy and the actions of policy
makers, goverments and organizations towards a sustainable future.

Derived from \citet{decclesia2024}: Among the ESG ETFs, the Clean Energy
(CE) ETFs have been the best-performing ones in 2022, followed by the
Cybersecurity and Artificial Intelligence (AI) ETFs. The clean energy
transition represents one of the largest multi-decade secular growth
opportunities.

From \citet{naqvi2022}: After the inclusion of Green energy financing in
the list of United Nations Sustainability Goals (SDGs) as SDG 7, the
role, importance, and visibility of green financial products; all have
escalated enormously.

@@

The sample consists of alternative energy equity ETFs (green ETFs) that
are traded in the US markets (see Table~\ref{tbl-data} in the
Appendix).\footnote{The data were sourced from
  \url{https://datastream.org/en-ca/}} The number of available
alternative energy ETFs in our sample varied from 10 in the beginning of
analysis to 30 at the most. The period of analysis runs from May
1\textsuperscript{st} of 2016 through 19\textsuperscript{th} June of
2024. The starting date was selected on the basis of the COP Paris
agreement. Daily logarithmic returns were computed from the closing
prices of ETFs for a total of 2122 observations.

@@

Following the relevant literature we calculate the Cross Sectional
Absolute Deviation (CSAD) measure for each day in the following manner.
We compute the difference of the ith ETF and market return where market
return is proxied by the cross sectional average of returns for sample
of our ETFs available for each day using Equation~\ref{eq-csad}:

\begin{equation}\phantomsection\label{eq-csad}{
CSAD_t = \dfrac{1}{N}\sum_{i=1}^{N} |R_{i,t} - R_{m,t}|
}\end{equation}

At a later stage we estimate a non-linear regression as in
\citet{galariotis2015}:

\begin{equation}\phantomsection\label{eq-herding}{
CSAD_t = \gamma_0 + \gamma_1 R_{m,t} + \gamma_2 R_{m,t}^2 + \epsilon_t
}\end{equation}

The testing procedure of herding relies on the above
Equation~\ref{eq-herding}. Rational asset pricing models predict a
linear relationship between return dispersion and market returns under
normal conditions, a relationship that is no longer valid in the
presence of herding. Herding behaviour leads to an increasing or
decreasing cross sectional dispersion with respect to market returns. In
other words, herding is captured by a non-linear term in the standard
pricing equation indicating a decreasing or an increasing returns'
dispersion.

@@

\begin{equation}\phantomsection\label{eq-asymmetry}{
CSAD_t = \gamma_0 + \gamma_1(1-D) R_{m,t} + \gamma_2D R_{m,t} + \gamma_3(1-D) R_{m,t}  + \gamma_4D R_{m,t}^2 \epsilon_t
}\end{equation}

where \(D\) is a dummy variable that takes the value of 1 when the
market return is negative and 0 otherwise. Therefore, our exploration of
asymmetric behavior of herding phenomenon is carried through the
inspection of the statistical significance and the sign of the two
estimated coefficients \(\gamma_3\) vs \(\gamma_4\) (up vs down
markets).

@@

\subsection{General herding behaviour}\label{general-herding-behaviour}

The behaviour of the CSAD measure for US Alternative Energy ETFs is
presented in Figure~\ref{fig-csad}.

@@

\begin{longtable}[l]{|p{1.39in}|p{1.39in}|p{1.39in}|p{1.39in}}

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Quantile}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\gamma_0$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\gamma_1 $}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\gamma_2$}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 10\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0016***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.2536***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.3736}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 25\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0026***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.2461***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.1056***}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 50\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0037***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.2648***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.165***}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 75\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0048***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.3011***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.1473***}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 90\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0064***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.2999***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.39in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ 0.2314}}}} \\

@@

\multicolumn{4}{>{\raggedright}m{\dimexpr 5.54in+6\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Note:\ This\ table\ presents\ the\ estimation\ results\ of\ herding\ of\ US\ Alternative\ energy\ equity\ ETFs\ according\ to\ Equation\ 2\ in\ various\ quantiles\ 10,\ 25,\ 50,\ 75\ and\ 90\%\ of\ the\ returns\ distribution.}}}\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{*,**,***\ denotes\ significance\ at\ 10\%,5\%\ and\ 1\%\ respectively.}}}} \\

@@

\subsection{Rolling window analysis}\label{rolling-window-analysis}

@@

\caption{\label{fig-rolling_window}Rolling window herding estimates}

@@

\subsection{Probit analysis}\label{probit-analysis}

The behaviour of participants in energy markets is closely related to
the developments in the field of climate risks, carbon emissions and
environmentally friendly policies. There are a few studies that attempt
to quantify the effects of uncertainty related to climate on the economy
and financial markets \textcolor{red}{(see inter alia….)}.
\citet{bua2024} developed two climate risk related indexes namely
transition and physical risk using a text-based approach in order to
study the effect of these risks in financial markets. It is expected
that investors would prefer to hold assets that perform well in the face
of increasing climate change risks, even if this entails accepting lower
returns for such climate-hedging assets. Therefore, in the context of
our study and following previous studies that study the determinants of
herding behaviour \citep[see][]{bouri2019, demirer2018}, we attempt to
study the effect of climate-related uncertainty on the formation of
herding behavior in the clean energy market. Given this, we define a
dummy variable, which takes a value of 1 during periods of statistically
significant herding (i.e., for days when the rolling t-statistic on
\(\gamma_2 < −1.96\)) and zero otherwise, and then, we use a Probit
model to relate this dummy to the two climate risk indexes developed by
\citet{bua2024}. It should be noted that due to availability issues the
probit analysis ends in December of 2023. The results from the Probit
model are reported in Table~\ref{tbl-probit_analysis}, where only the
physical risk index significantly decreases the probability of herding.
In other words, climate risks is good news for clean energy stocks or
firms resulting in anti herding behaviour.

This implies that in the presence of higher physical risk with respect
to the climate, clean energy ETFs become a more attractive investment
option for investors that allocate their money to the various
alternative energy investment products. As a result, the cross sectional
dispersion of clean energy ETFs tends to increase.

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Transition\ Risk}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 2.14in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{-4.607***}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Physical\ Risk}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 2.14in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{-1.318}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Constant}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 2.14in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.506***}}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{LogL}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 2.14in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{-484.7}}}} \\

@@

\begin{longtable}[l]{|p{2.23in}|p{1.90in}|p{1.86in}}

@@

@@

\subsection{Data}\label{data-1}

>>>>

\author { 
Vassilios Babalos\footnote{Corresponding Author. Department of Accounting and Finance, University of Peloponnese, Kalamata,  Greece; Email: v.babalos@uop.gr.} \,\,
Xolani Sibande\footnote{Department of Economics, University of Pretoria, Pretoria, South Africa; Email: xolaniss@gmail.com. Economic Research Department, South African Reserve Bank, Pretoria, South Africa.} \,\, 
Elie Bouri\footnote{Adnan Kassar School of Business, Lebanese American University, Lebanon; Email: elie.elbouri@lau.edu.lb} \,\,
Rangan Gupta\footnote{Department of Economics, University of Pretoria, Pretoria, South Africa; Email: rangan.gupta@up.ac.za.} 
}
\date{\today}
\maketitle

@@

Investment interest on green ETFs has been explosive mainly by climate
risks faced by the global economy and the actions of policy makers,
governments and organizations towards a sustainable future. Among the
ESG ETFs, the Clean Energy (CE) ETFs have been the best-performing ones
in 2022, followed by the Cybersecurity and Artificial Intelligence (AI)
ETFs \citep{decclesia2024}. The clean energy transition represents one
of the largest multi-decade secular growth opportunities. After the
inclusion of Green energy financing in the list of United Nations
Sustainability Goals (SDGs) as SDG 7, the role, importance, and
visibility of green financial products; all have escalated enormously
\citep{naqvi2022}.

@@

The sample consists of alternative energy equity ETFs (green ETFs) that
are traded in the US markets (see Table~\ref{tbl-data} in the
Appendix).\footnote{The data were sourced from
  \url{https://datastream.org/en-ca/}} The number of available
alternative energy ETFs in our sample varied from 10 in the beginning of
analysis to 30 at the most. The period of analysis runs from May
1\textsuperscript{st} of 2016 through 19\textsuperscript{th} June of
2024. The starting date was selected on the basis of the UN Climate
Change Conference (COP) Paris agreement. Daily logarithmic returns were
computed from the closing prices of ETFs for a total of 2122
observations.

@@

Following the relevant literature
\citetext{\citealp{christie1995}; \citealp[and][]{chang2000}}, we
compute dispersion of the \(ith\) ETF from the market return. This is
known as the Cross Sectional Absolute Deviation (\(CSAD_t\)) measure.
Empirically the \(CSAD_t\) is defined in the following manner:

\begin{equation}\phantomsection\label{eq-csad}{
CSAD_t = \dfrac{1}{N}\sum_{i=1}^{N} |R_{i,t} - R_{m,t}|,
}\end{equation}

where \(R_{i,t}\) is the return and \(R_{m,t}\) is the cross sectional
average of returns for the sample of ETFs available for each day. The
return dispersion measures the directional similarity of ETF returns to
the market return. This return similarity forms the basis for the
herding behaviour tests. The CSAD measure for US Alternative Energy ETFs
is presented in Figure~\ref{fig-csad}. Following \citet{galariotis2015}
we estimate Equation~\ref{eq-herding}:

\begin{equation}\phantomsection\label{eq-herding}{
CSAD_t = \gamma_0 + \gamma_1 R_{m,t} + \gamma_2 R_{m,t}^2 + \epsilon_t,
}\end{equation}

where \(\gamma_0\) is the intercept, \(\gamma_1\) is the coefficient of
the linear term, \(\gamma_2\) is the coefficient of the quadratic term
or the herding behaviour term, and \(\epsilon_t\) is the error term. The
coefficient \(\gamma_2 <0\) when herding is present, and \(\gamma_2 >0\)
when anti-herding is present. To ensure the robustness of the estimate,
we estimate \(CSAD_{t}\) with Newey-West standard errors
\citep[See][]{newey1987}.

@@

\begin{equation}\phantomsection\label{eq-asymmetry}{
CSAD_t = \gamma_0 + \gamma_1(1-D) R_{m,t} + \gamma_2D R_{m,t} + \gamma_3(1-D) R_{m,t}^2  + \gamma_4D R_{m,t}^2 + \epsilon_t,
}\end{equation}

where \(D\) is a dummy variable that takes the value of 1 when the
market return is negative and 0 otherwise. Therefore, our exploration of
asymmetric behaviour of herding phenomenon is carried through the
inspection of the statistical significance and the sign of the two
estimated coefficients \(\gamma_3\) versus \(\gamma_4\) (up versus down
markets).

@@

\subsection{Herding behaviour}\label{herding-behaviour}

Rational asset pricing models \citep[for example,][]{black1972} predict
a linear relationship between return dispersion and market returns under
normal conditions, a relationship that is no longer valid in the
presence of herding. Herding behaviour leads to an increasing or
decreasing cross sectional dispersion with respect to market returns. In
other words, herding is captured by a non-linear term in the standard
pricing equation indicating a decreasing or an increasing returns'
dispersion.

@@

\begin{longtable}[l]{|p{1.49in}|p{1.49in}|p{1.49in}|p{1.49in}}

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Quantile}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\gamma_0$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\gamma_1 $}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\gamma_2$}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 10\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0016***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.2536***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.3736}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 25\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0026***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.2461***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.1056***}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 50\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0037***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.2648***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.165***}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 75\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0048***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.3011***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.1473***}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\tau = 90\%$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.0064***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{0.2999***}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.49in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ 0.2314}}}} \\

@@

\multicolumn{4}{>{\raggedright}m{\dimexpr 5.94in+6\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Note:\ This\ table\ presents\ the\ estimation\ results\ of\ herding\ of\ US\ Alternative\ energy\ equity\ ETFs\ according\ to\ Equation\ 2\ in\ various\ quantiles\ 10,\ 25,\ 50,\ 75\ and\ 90\%\ of\ the\ returns\ distribution.}}}\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{*,**,***\ denotes\ significance\ at\ 10\%,5\%\ and\ 1\%\ respectively.}}}} \\

@@

\subsection{Herding behaviour
over-time}\label{herding-behaviour-over-time}

@@

\caption{\label{fig-rolling_window}Rolling window herding estimates.
Note: The red perforated lines indicates the 95\% confidence interval.}

@@

\subsection{Climate-related uncertainty and herding
behaviour}\label{climate-related-uncertainty-and-herding-behaviour}

The behaviour of participants in energy markets is closely related to
the developments in the field of climate risks, carbon emissions and
environmentally friendly policies. There are a few studies that attempt
to quantify the effects of uncertainty related to climate on the economy
and financial markets \textcolor{red}{(see inter alia….)}.
\citet{bua2024} developed two climate risk related indexes namely
transition and physical risk using a text-based approach in order to
study the effect of these risks in financial markets. It is expected
that investors would prefer to hold assets that perform well in the face
of increasing climate change risks, even if this entails accepting lower
returns for such climate-hedging assets. Therefore, in the context of
our study and following previous studies that study the determinants of
herding behaviour \citep[see][]{bouri2019, demirer2018}, we attempt to
study the effect of climate-related uncertainty on the formation of
herding behavior in the clean energy market.

We use a probit model to relate herding to the two climate risk indexes
developed by \citet{bua2024} in the following manner:

\begin{equation}\phantomsection\label{eq-climate}{
Pr(D^{herd} = 1 |\lambda_0 +  \lambda_1TRI  + \lambda_2PRI) = \lambda_0 +  \lambda_1TRI  + \lambda_2PRI,
}\end{equation}

where \(D^{herd}\) takes a value of 1 during periods of statistically
significant herding (i.e., for days when the rolling t-statistic on
\(\gamma_2 < −1.96\) in Figure~\ref{fig-rolling_window}) and zero
otherwise. \(TRI\) is the transitional risk index and \(PRI\) is the
physical risk index.

The results from the Probit model are reported in
Table~\ref{tbl-probit_analysis}, where only the physical risk index
significantly decreases the probability of herding.\footnote{It should
  be noted that due to availability issues the probit analysis ends in
  December of 2023.} In other words, climate risks is good news for
clean energy stocks or firms resulting in anti-herding behaviour. This
implies that in the presence of higher physical risk with respect to the
climate, clean energy ETFs become a more attractive investment option
for investors that allocate their money to the various alternative
energy investment products. As a result, the cross sectional dispersion
of clean energy ETFs tends to increase.

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\lambda_0$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 2.14in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ -1.506***}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\lambda_1$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 2.14in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{-4.607***}}}} \\

@@

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{$\lambda_2$}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 2.14in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{-1.318}}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Log\ Likelihood}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 2.14in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{-484.7}}}} \\

@@

\begin{longtable}[l]{|p{2.59in}|p{1.90in}|p{1.86in}}

@@

We use these high PRI and high TRI in one probit regression and low TRI
and low PRI in another. Results are presented
Table~\ref{tbl-probit_analysis2}. We observe higher uncertainty that
stems from physical or transition risk causes anti-herding which is in
line with the logic we discussed earlier.

@@

\subsection{Cross-sectional average
deviations}\label{cross-sectional-average-deviations}
